Drug sensitivity prediction with normal inverse Gaussian shrinkage informed by external data

### 2020-01-30 docs/manuscript1.Rnw
added Bayesian ridge and lasso

	\subsection{Bayesian prior model}
	We carry out inference by endowing the parameters with the following priors:
	\begin{subequations}\label{eq:prior}
		\begin{align}
		  \beta_{jd} | \gamma^2_{jd}, \tau_d^2, \sigma^2_d & \sim \mathcal{N}_{p} 
		  (0, \gamma_{jd}^2 \tau_d^2 \sigma_d^2), 
		  \label{eq:betaprior}\\
		  \gamma_{jd}^{2} & \sim \mathcal{IG}(\phi_{jd}, \lambda_{\text{feat}}), \\
		  \tau_d^{2} & \sim \mathcal{IG}(\chi_d, \lambda_{\text{drug}}), 
		  \label{eq:tauprior}\\
		  \sigma_d^{2} & \sim 1/\sigma_d^{3},
		\end{align}
	\end{subequations}
	where $\mathcal{IG}(\phi, \lambda)$ denotes an inverse Gaussian distribution 
	with mean $\phi$ and shape $\lambda$. Prior distributions of the form
	(\ref{eq:prior}) are often referred to as global-local shrinkage rules 
	\cite[]{bernardo_shrink_2011}, due to the multiplicative separation of the 
	prior 
	variance into a local component $\gamma_{jd}^2$ and a global component
	$\tau_d^2$. The local components are supposed to capture local variation in
	the model parameters $\beta_{jd}$, while the global component is supposed to
	capture the general trend in $\bm{\beta}_{d}$. For appropriate 
	local shrinkage in global-local 
	shrinkage models it is important  to account for 
  different noise levels $\sigma_d^2$ by scaling the $\beta_{jd}$ variances 
  accordingly.

@@

tab <- cbind(apply(res[rownames(res)=="emsel", ], 2, median),
             apply(res[rownames(res)=="emseh", ], 2, median),
             apply(res[rownames(res)=="emse", ], 2, median),
             apply(res[rownames(res)=="pmse", ], 2, median))
tab <- round(tab, 3)
tab[which(tab[, 1]==min(tab[, 1], na.rm=TRUE)), 1] <- 
  paste0("\\textbf{", tab[which(tab[, 1]==min(tab[, 1], na.rm=TRUE)), 1], "}")
tab[which(tab[, 2]==min(tab[, 2], na.rm=TRUE)), 2] <- 
  paste0("\\textbf{", tab[which(tab[, 2]==min(tab[, 2], na.rm=TRUE)), 2], "}")
tab[which(tab[, 3]==min(tab[, 3], na.rm=TRUE)), 3] <- 
  paste0("\\textbf{", tab[which(tab[, 3]==min(tab[, 3], na.rm=TRUE)), 3], "}")
tab[which(tab[, 4]==min(tab[, 4], na.rm=TRUE)), 4] <- 
  paste0("\\textbf{", tab[which(tab[, 4]==min(tab[, 4], na.rm=TRUE)), 4], "}")
colnames(tab) <- c("$\\text{EMSE}_{\\text{bottom}}$",
                   "$\\text{EMSE}_{\\text{top}}$",
                   "$\\text{EMSE}_{\\text{total}}$",
                   "$\\text{PMSE}$")     
rownames(tab)[c(1, 2)] <- c("NIG$_{\\text{f}+\\text{d}}^-$",
                            "NIG$_{\\text{f}+\\text{d}}$")
@	
  In Figure \ref{fig:simulations_gdsc_est3} the 
	$\bm{\alpha}_{\text{feat}}$ and $\bm{\alpha}_{\text{drug}}$ estimated by 
	NIG$_{\text{f}+\text{d}}$
	are displayed together with their true values for
	simulation 3. $\bm{\alpha}_{\text{feat}}$ are underestimated, while 
	$\bm{\alpha}_{\text{drug}}$ are overestimated. These biases seem to 
	be consistent though. The median ratios $\phi_1$ to $\phi_2$, $\phi_3$, 
	$\phi_4$ and $\chi_1$ to $\chi_2$, $\chi_3$, 
	$\chi_4$ are \Sexpr{round(phi.ratio, 2)[-1]} and 
	\Sexpr{round(chi.ratio, 2)[-1]}, respectively, compared to their true values
	\Sexpr{round(true.ratio, 2)[-1]}. So in a relative sense, the estimates are
	correct. Median $\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ are
	are over- and underestimated with \Sexpr{round(est.lambdaf[2], 2)}
	and \Sexpr{round(est.lambdad[2], 2)}, respectively, compared to their true
	value 1. The NIG$_{\text{f}+\text{d}}^-$ model is also consistently over- 
	and underestimating $\phi$
	and $\chi$ with estimates \Sexpr{round(est.phi, 2)} and 
	\Sexpr{round(est.chi, 2)}, respectively (compared to the true mean 
	\Sexpr{round(true.chi, 2)}). NIG$_{\text{f}+\text{d}}^-$ also over- and
	underestimates 
	$\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ with median estimates
	\Sexpr{round(est.lambdaf[1], 2)}
	and \Sexpr{round(est.lambdad[1], 2)}, respectively. Table 
	\ref{tab:simulations_gdsc_tab3}, with median performance measures, shows
	that NIG$_{\text{f}+\text{d}}$ outperforms the other models in terms of
	prediction and estimation MSE.
<<simulations_gdsc_est3, fig.cap="Estimated (a) hyperparameters $\\bm{\\alpha}_{\\text{feat}}$ and (b) prior means $\\bm{\\phi}_{jd}$ from simulation version 3.", out.width="100%", fig.asp=1>>= 
@

<<simulations_gdsc_tab3>>=
library(knitr, quietly=TRUE, verbose=FALSE)
suppressWarnings(library(kableExtra))
options(knitr.kable.NA="", digits=3)
kable(tab, col.names=colnames(tab), align="r",
      caption="Median performance measure for simulation version 3, estimated on the test data (lowest in bold).", 
      format="latex", booktabs=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("HOLD_position"))
@

@@

	A comprehensive penalized regression
	approach such as in 
	{\cite{mai_composite_2019} uses most of the available molecular markers 
	instead of just gene expression (68 mutations, 426 copy numbers, 
	and 2602 gene expression levels for 498 cell lines and 97 drugs). The 
	current version of our software is not able to handle such problem sizes,
	due to the repeated calculation of posterior parameters in the empirical
	Bayes iterations. Currently, problems with number of outcomes in the hundreds
	and number of features and samples in the thousands are computationally 
	feasible.

@@

	To compute performance measures, we randomly split the data into training and
	test sets, containing 
	241 cell lines each, estimate the models on the training data and compute 
	performance measures on the test data. We repeated this procedure 50 times
	and average performance measures over the splits.
	The estimated 
	models are: four models like in (\ref{eq:prior}) but with different external 
	variables, named NIG$_{\text{f}+\text{d}}^-$, NIG$_{\text{f}+\text{d}}$,
	NIG$_{\text{f}}$, and NIG$_{\text{d}}$ in the following. The nomenclature is
	the same as in Section \ref{sec:simulations}, so NIG$_{\text{f}+\text{d}}^-$
	does not use the external covariates and estimates
	one overall $\phi$ and one overall $\chi$, while NIG$_{\text{f}+\text{d}}$ 
	uses the external covariates to estimate both $\bm{\alpha}_{\text{feat}}$ and
	$\bm{\alpha}_{\text{drug}}$. NIG$_{\text{f}}$ and NIG$_{\text{d}}$ estimate 
	only $\bm{\alpha}_{\text{feat}}$ or $\bm{\alpha}_{\text{drug}}$, respectively,
	keeping either $\forall d: \tau_d^2=1$ or $\forall j,d: \gamma_{jd}^2=1$
	fixed. In the language of \cite{bernardo_shrink_2011} as introduced in 
	Section \ref{sec:introduction}, models NIG$_{\text{f}}$ and NIG$_{\text{d}}$
	may be described as local and global shrinkage rules, respectively, as 
	opposed to the global-local shrinkage models NIG$_{\text{f}+\text{d}}$ and
	NIG$_{\text{f}+\text{d}}^-$.

	In addition, we estimate regular lasso and ridge models, and a
	bSEM model \cite[]{leday_gene_2017,kpogbezan_empirical_2017}. The bSEM model 
	is similar to the NIG model in that it
	draws the $\beta_{jd}$ from a conditionally normal distribution, but
	fixes $\tau_d^2=1$ and draws
	$\gamma_{jd}^2 \sim \Gamma^{-1} (\phi_{jd}, \lambda_{jd})$, instead of
	$\gamma_{jd}^2 \sim \mathcal{IG} (\phi_{jd}, \lambda_{jd})$. Additionally,
	the bSEM model can
	only incorporate one binary external covariate that divides the features into
	two groups, i.e., $\bm{\alpha}_{\text{feat}} = \begin{bmatrix} \alpha_0 &
	\alpha_1
	\end{bmatrix} \tr$ and $\bm{c}_{jd} = \begin{bmatrix} 1 & 0 \end{bmatrix}
	\tr$ if feature $jd$ belongs to group one and $\bm{c}_{jd} = \begin{bmatrix}
	1 & 1 \end{bmatrix} \tr$ if feature $jd$ belongs to group two. Both
	$\bm{\alpha}_{\text{feat}}$ and $\lambda_{\text{feat}}$ are
	estimated in a similar fashion as in Section \ref{sec:estimation}.
	The null model that fits the means of the $\y_d$ is included as a reference.

@@

tab <- cbind(apply(res[rownames(res)=="pmse", ], 2, median))
tabm <- tab
tabm <- round(tabm, 3)
tabm[which(tabm[, 1]==min(tabm[, 1], na.rm=TRUE)), 1] <- 
  paste0("\\textbf{", tabm[which(tabm[, 1]==min(tabm[, 1], na.rm=TRUE)), 1], "}")
colnames(tabm) <- c("$\\text{PMSE}_{\\text{test}}$")

library(glmnet)
kurtosis <- function(x) {
  n <- length(x) 
  kurt <- (n + 1)*n*sum((x - mean(x))^4)/((n - 1)*(n - 2)*(n - 3)*var(x)^2) - 
    3*(n - 1)^2/((n - 2)*(n - 3))
  replace(kurt, !is.finite(kurt), NA)
}
load("results/analysis_gdsc_fit1.Rdata")
best2.semnig <- fit2.semnig$vb$mu
best1.lasso <- lapply(fit1.lasso, function(s) {
  unname(coef(s, "lambda.min")[-1, ])})
best1.ridge <- lapply(fit1.ridge, function(s) {
  unname(coef(s, "lambda.min")[-1, ])})
best1.bSEM <- lapply(fit1.bSEM$vb$beta, function(s) {s[, 2]})
kurts <- c(mean(sapply(best2.semnig, kurtosis), na.rm=TRUE), 
           mean(sapply(best1.lasso, kurtosis), na.rm=TRUE),
           mean(sapply(best1.ridge, kurtosis), na.rm=TRUE), 
           mean(sapply(best1.bSEM, kurtosis), na.rm=TRUE))
ranges <- rbind(range(unlist(lapply(1:length(
  fit2.semnig$eb$mpriord), function(d) {
    fit2.semnig$eb$mpriord[d]*fit2.semnig$eb$mpriorf[[d]]}))),
  fit1.bSEM$eb$b[nrow(fit1.bSEM$eb$b), ]/
    (fit1.bSEM$eb$a[nrow(fit1.bSEM$eb$a), ] - 1))
@
  The non-zero $\bm{\alpha}$ estimates in Tables 
  \ref{tab:analysis_gdsc_est1_feat} and \ref{tab:analysis_gdsc_est1_drug} show 
  that there is an effect of the
  external covariates. Models NIG$_{\text{f}+\text{d}}$, NIG$_{\text{f}}$, and
  bSEM estimate a negative effect for
  the external pathway covariate. This translates to a negative additive effect 
  on the expected prior precision of the $\bm{\beta}_d$, i.e., genes that are 
  found
  in the target pathway of a drug are more predictive of IC50 than genes that
  are not in the target pathway, according to expectation. Furthermore the 
  NIG$_{\text{f}+\text{d}}$
  model estimates positive effects for the experimental and developmental drugs,
  compared to the approved drugs. This translates to smaller expected 
  prior precisions for the regression parameters for approved drugs, i.e.,
  approved drugs are easier to predict than experimental and developmental 
  drugs. On the other hand, the NIG$_{\text{d}}$ model, that only includes
  external
  drug covariates, estimates a negative effect for the developmental drugs.
  The NIG$_{\text{f}+\text{d}}$ model estimates the smallest prior 
  $\bm{\beta}_d$ precision for the
  targeted drugs, followed by cytotoxic drugs, while drugs with an unkown target
  are assigned the largest prior $\bm{\beta}_d$ precision, according
  to expectation. Somewhat surprisingly, the NIG$_{\text{d}}$ model
  estimates the smallest
  prior $\bm{\beta}_d$ precision for the cytotoxic drugs, compared to the 
  targeted and unkown drugs. We conjecture that, due to absence of a local
  feature-specific variance component, this model is not flexible 
  enough to correctly capture the full covariance structure of the 
  $\bm{\beta}_d$ and therefore incorreclty estimates $\bm{\alpha}$.
  To see the total effect of the external covariates on the $\bm{\beta}_d$ 
  prior, we calculated the range of the expected prior variances 
  (up to a multiplicative error variance) as 
  $\hat{E}(\gamma_{jd}^2 \cdot \tau_d^2) \in \[ \Sexpr{ranges[1, ]}\]$ and
  $\hat{E}(\gamma_{jd}^2 \cdot \tau_d^2) \in \[ \Sexpr{ranges[2, ]}\]$ for
  NIG$_{\text{f}+\text{d}}$ and bSEM, respectively. This confirms that the
  estimated prior 
  $\bm{\beta}_d$ variance components are in the same order of magnitude and that
  the external covariates have a detectable effect on the prior.
<<analysis_gdsc_est1_feat>>=
options(knitr.kable.NA="", digits=3)
kable(est1,
      col.names=colnames(est1), align="r",
      caption="$\\bm{\\alpha}_{\\text{feat}}$ estimates from GDSC data.",
      format="latex", booktabs=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("HOLD_position"))
@

<<analysis_gdsc_est1_drug>>=
kable(est2,
      col.names=colnames(est2), align="r",
      caption="$\\bm{\\alpha}_{\\text{drug}}$ estimates from GDSC data.",
      format="latex", booktabs=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("HOLD_position"))
@
  The median PMSEs and ELBOs, calculated on the test data
  are displayed in Table \ref{tab:analysis_gdsc_tab1}. In terms of prediction 
  error, ridge outperforms the other models 
  (PMSE of \Sexpr{round(tab[6, 1], 3)}), while bSEM 
  (PMSE of \Sexpr{round(tab[7, 1], 3)}) performs worse than 
  the null model (PMSE of \Sexpr{round(tab[8, 1], 3)}). Second and third best
  performing models are NIG$_{\text{f}+\text{d}}^-$ and 
  NIG$_{\text{f}+\text{d}}$ (PMSEs of \Sexpr{round(tab[1, 1], 3)} and
  \Sexpr{round(tab[2, 1], 3)}, respectively), both outperforming the lasso
  \Sexpr{round(tab[5, 1], 3)}. The underperformance of the 
  NIG$_{\text{f}}$ and NIG$_{\text{d}}$ models shows the usefulness of a
  global-local shrinkage prior as opposed to only global or local shrinkage as 
  as in models NIG$_{\text{f}}$ and NIG$_{\text{d}}$.
<<analysis_gdsc_tab1>>=
kable(tabm,
      col.names=colnames(tabm), align="r",
      caption="Performance measures calculated on GDSC data (best performing model in bold).",
      format="latex", booktabs=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("HOLD_position"))
@

@@

	Simulations in Section \ref{sec:simulations} show that the method is 
	reasonably able to recover hyperparameters. Simulation versions 1 and 2 show
	that estimation of drug- and feature-specific hyperparameters is, in 
	principle, fairly accurate. However, when estimated jointly, biases may occur
	due to the interplay between the two sources of information. Nevertheless,
	the simulations show that inclusion of informative external covariates is 
	still beneficial to predictive performance.

	The model is put into practice on the GDSC data. 
	The NIG model is competitive with convential models like lasso and ridge.
	We note that estimation of the penalty parameters in the latter two models
	is geared towards predictive performance, because cross-validation is used
	to train these parameters, as opposed to estimation of the
	hyperparameters in the NIG model, which aims to improve model fit. In that 
	sense, the comparison metric (PMSE) is not to the benefit of the NIG model.

@@

	A possible direction of future research that attempts 
	to lower PMSE of the NIG model is the introduction of a cross-validated global
	variance component in the prior of the $\bm{\beta}_d$. If we add the 
	additional constraint that the drug and feature variance components
	average to one, this cross-validated variance component may be interpreted as
	the overall level of shrinkage. This overall shrinkage level is then tailored 
	to lower PMSE. Other directions of future research are applications of the NIG
	model to different data types. An example of such an application is eQTL, 
	where gene expressions are regressed on SNPs. Several interesting external
	covariates are available, both on the genes as well as the SNPs. An example of
	an external
	covariate for the genes is genelength, where we suspect that longer genes are
	harder to predict. The distance of the SNP to the gene is an example of an
	external SNP covariate, where the expectation is that SNPs further from the 
	gene are less predictive of that gene's expression.

>>>>

	\subsection{Bayesian prior model}\label{sec:prior}
	We carry out inference by endowing the parameters with the following priors:
	\begin{subequations}\label{eq:prior}
		\begin{align}
		  \beta_{jd} | \gamma^2_{jd}, \tau_d^2, \sigma^2_d & \sim \mathcal{N}_{p} 
		  (0, \gamma_{jd}^2 \tau_d^2 \sigma_d^2), 
		  \label{eq:betaprior}\\
		  \gamma_{jd}^{2} & \sim \mathcal{IG}(\phi_{jd}, \lambda_{\text{feat}}), \\
		  \tau_d^{2} & \sim \mathcal{IG}(\chi_d, \lambda_{\text{drug}}), 
		  \label{eq:tauprior}\\
		  \sigma_d^{2} & \sim 1/\sigma_d^{3},
		\end{align}
	\end{subequations}
	where $\mathcal{IG}(\phi, \lambda)$ denotes an inverse Gaussian distribution 
	with mean $\phi$ and shape $\lambda$.

	In model (\ref{eq:prior}), $\gamma_{jd}^2$ denotes a local variance 
	component that is supposed to capture local, feature-specific variation in the
	model parameters $\beta_{jd}$, while the global variance component $\tau$ 
	captures the, drug-specific, general trend in $\bm{\beta}_{d}$. Each drug is
	endowed with a random error variance $\sigma_d^2$.

	Prior distributions of the form
	(\ref{eq:prior}) are often referred to as global-local shrinkage rules 
	\cite[]{bernardo_shrink_2011}, due to the multiplicative separation of the 
	prior variance into a local component $\gamma_{jd}^2$ and a global component
	$\tau_d^2$. For appropriate local shrinkage in global-local 
	shrinkage models it is important  to account for 
  different noise levels $\sigma_d^2$ by scaling the $\beta_{jd}$ variances 
  accordingly.

@@

tab <- cbind(apply(res[rownames(res)=="emsel", ], 2, median),
             apply(res[rownames(res)=="emseh", ], 2, median),
             apply(res[rownames(res)=="emse", ], 2, median),
             apply(res[rownames(res)=="pmse", ], 2, median))
tab <- round(tab, 3)
tab[which(tab[, 1]==min(tab[, 1], na.rm=TRUE)), 1] <- 
  paste0("\\textbf{", tab[which(tab[, 1]==min(tab[, 1], na.rm=TRUE)), 1], "}")
tab[which(tab[, 2]==min(tab[, 2], na.rm=TRUE)), 2] <- 
  paste0("\\textbf{", tab[which(tab[, 2]==min(tab[, 2], na.rm=TRUE)), 2], "}")
tab[which(tab[, 3]==min(tab[, 3], na.rm=TRUE)), 3] <- 
  paste0("\\textbf{", tab[which(tab[, 3]==min(tab[, 3], na.rm=TRUE)), 3], "}")
tab[which(tab[, 4]==min(tab[, 4], na.rm=TRUE)), 4] <- 
  paste0("\\textbf{", tab[which(tab[, 4]==min(tab[, 4], na.rm=TRUE)), 4], "}")
colnames(tab) <- c("$\\text{EMSE}_{\\text{bottom}}$",
                   "$\\text{EMSE}_{\\text{top}}$",
                   "$\\text{EMSE}_{\\text{total}}$",
                   "$\\text{PMSE}$")     
rownames(tab)[c(1, 2)] <- c("NIG$_{\\text{f}+\\text{d}}^-$",
                            "NIG$_{\\text{f}+\\text{d}}$")
@	
  In Figure \ref{fig:simulations_gdsc_est3} the 
	$\bm{\alpha}_{\text{feat}}$ and $\bm{\alpha}_{\text{drug}}$ estimated by 
	NIG$_{\text{f}+\text{d}}$
	are displayed together with their true values for
	simulation Scenario 3. $\bm{\alpha}_{\text{feat}}$ are underestimated, while 
	$\bm{\alpha}_{\text{drug}}$ are overestimated. These biases seem to 
	be consistent though. The median ratios $\phi_1$ to $\phi_2$, $\phi_3$, 
	$\phi_4$ and $\chi_1$ to $\chi_2$, $\chi_3$, 
	$\chi_4$ are \Sexpr{round(phi.ratio, 2)[-1]} and 
	\Sexpr{round(chi.ratio, 2)[-1]}, respectively, compared to their true values
	\Sexpr{round(true.ratio, 2)[-1]}. So in a relative sense, the estimates are
	correct. Median $\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ are
	are over- and underestimated with \Sexpr{round(est.lambdaf[2], 2)}
	and \Sexpr{round(est.lambdad[2], 2)}, respectively, compared to their true
	value 1. The NIG$_{\text{f}+\text{d}}^-$ model is also consistently over- 
	and underestimating $\phi$
	and $\chi$ with estimates \Sexpr{round(est.phi, 2)} and 
	\Sexpr{round(est.chi, 2)}, respectively (compared to the true mean 
	\Sexpr{round(true.chi, 2)}). NIG$_{\text{f}+\text{d}}^-$ also over- and
	underestimates 
	$\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ with median estimates
	\Sexpr{round(est.lambdaf[1], 2)}
	and \Sexpr{round(est.lambdad[1], 2)}, respectively. Table 
	\ref{tab:simulations_gdsc_tab3}, with median performance measures, shows
	that NIG$_{\text{f}+\text{d}}$ outperforms the other models in terms of
	prediction and estimation MSE, with again most of the edge over ridge in the
	top 10\% of the $\beta_{jd}$.
<<simulations_gdsc_est3, fig.cap="Simulation results for Scenario 3: (a) estimated and true values for $\\alpha_{\\text{feat}}$ and, (b) estimated and true values for prior means $\\phi_{jd}$.", out.width="100%", fig.asp=1>>= 
@

<<simulations_gdsc_tab3>>=
library(knitr, quietly=TRUE, verbose=FALSE)
suppressWarnings(library(kableExtra))
options(knitr.kable.NA="", digits=3)
kable(tab, col.names=colnames(tab), align="r",
      caption="Median performance measure for simulation Scenario 3, estimated on the test data (lowest in bold).", 
      format="latex", booktabs=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("HOLD_position"))
@

@@

	A comprehensive penalized regression
	approach such as in 
	{\cite{mai_composite_2019} uses most of the available molecular markers 
	instead of just gene expression (68 mutations, 426 copy numbers, 
	and 2602 gene expression levels for 498 cell lines and 97 drugs). The 
	current version of our software is not able to handle such problem sizes,
	due to the repeated calculation of posterior parameters in the empirical
	Bayes iterations. Currently, problems with number of outcomes in the hundreds
	and number of features and samples up to a thousand are computationally 
	feasible.

@@

	To compute performance measures, we randomly split the data into training and
	test sets, containing 
	241 cell lines each, estimate the models on the training data and compute 
	performance measures on the test data. We repeated this procedure 50 times
	and average performance measures over the splits.
	The estimated 
	models are: four models like in (\ref{eq:prior}) but with different external 
	variables, named NIG$_{\text{f}+\text{d}}^-$, NIG$_{\text{f}+\text{d}}$,
	NIG$_{\text{f}}$, and NIG$_{\text{d}}$ in the following. The nomenclature is
	the same as in Section \ref{sec:simulations}, i.e., 
	NIG$_{\text{f}+\text{d}}^-$ excludes the external covariates and estimates
	one overall $\phi$ and one overall $\chi$, while NIG$_{\text{f}+\text{d}}$ 
	uses the external covariates to estimate both $\bm{\alpha}_{\text{feat}}$ and
	$\bm{\alpha}_{\text{drug}}$. NIG$_{\text{f}}$ and NIG$_{\text{d}}$ estimate 
	only $\bm{\alpha}_{\text{feat}}$ or $\bm{\alpha}_{\text{drug}}$, respectively,
	keeping either $\forall d: \tau_d^2=1$ or $\forall j,d: \gamma_{jd}^2=1$
	fixed. In the language of \cite{bernardo_shrink_2011} as introduced in 
	Section \ref{sec:introduction}, models NIG$_{\text{f}}$ and NIG$_{\text{d}}$
	may be described as local and global shrinkage rules, respectively, as 
	opposed to the global-local shrinkage models NIG$_{\text{f}+\text{d}}$ and
	NIG$_{\text{f}+\text{d}}^-$.

	In addition, we estimate regular lasso and ridge models, and a
	bSEM model \cite[]{leday_gene_2017,kpogbezan_empirical_2017}. The bSEM model 
	is similar to the NIG model in that it
	draws the $\beta_{jd}$ from a conditionally normal distribution, but
	fixes $\tau_d^2=1$ and draws
	$\gamma_{jd}^2 \sim \Gamma^{-1} (\phi_{jd}, \lambda_{jd})$, instead of
	$\gamma_{jd}^2 \sim \mathcal{IG} (\phi_{jd}, \lambda_{jd})$, where 
	$\Gamma^{-1} (\phi, \lambda)$ denotes the inverse gamma distribution with 
	shape $\phi$ and scale $\lambda$. Additionally, the bSEM model can
	only incorporate one binary external covariate that divides the features into
	two groups, i.e., $\bm{\alpha}_{\text{feat}} = \begin{bmatrix} \alpha_0 &
	\alpha_1
	\end{bmatrix} \tr$ and $\bm{c}_{jd} = \begin{bmatrix} 1 & 0 \end{bmatrix}
	\tr$ if feature $jd$ belongs to group one and $\bm{c}_{jd} = \begin{bmatrix}
	1 & 1 \end{bmatrix} \tr$ if feature $jd$ belongs to group two. Both
	$\bm{\alpha}_{\text{feat}}$ and $\lambda_{\text{feat}}$ are
	estimated in a similar fashion as in Section \ref{sec:estimation}.
	The null model that fits the means of the $\y_d$ is included as a reference.

@@

rownames(est1)[c(1:4)] <- rownames(est2)[c(1:4)] <-
  c("NIG$_{\\text{f}+\\text{d}}^-$", "NIG$_{\\text{f}+\\text{d}}$",
    "NIG$_{\\text{f}}$", "NIG$_{\\text{d}}$")

tab <- cbind(apply(res[rownames(res)=="pmse", ], 2, median))
tabm <- tab
tabm <- round(tabm, 3)
tabm[which(tabm[, 1]==min(tabm[, 1], na.rm=TRUE)), 1] <- 
  paste0("\\textbf{", tabm[which(tabm[, 1]==min(tabm[, 1], na.rm=TRUE)), 1], "}")
colnames(tabm) <- c("$\\text{PMSE}_{\\text{test}}$")  
rownames(tabm)[c(1:4)] <- rownames(est2)[c(1:4)] <-
  c("NIG$_{\\text{f}+\\text{d}}^-$", "NIG$_{\\text{f}+\\text{d}}$",
    "NIG$_{\\text{f}}$", "NIG$_{\\text{d}}$")

library(glmnet)
kurtosis <- function(x) {
  n <- length(x) 
  kurt <- (n + 1)*n*sum((x - mean(x))^4)/((n - 1)*(n - 2)*(n - 3)*var(x)^2) - 
    3*(n - 1)^2/((n - 2)*(n - 3))
  replace(kurt, !is.finite(kurt), NA)
}
load("results/analysis_gdsc_fit1.Rdata")
best2.semnig <- fit2.semnig$vb$mu
best1.lasso <- lapply(fit1.lasso, function(s) {
  unname(coef(s, "lambda.min")[-1, ])})
best1.ridge <- lapply(fit1.ridge, function(s) {
  unname(coef(s, "lambda.min")[-1, ])})
best1.bSEM <- lapply(fit1.bSEM$vb$beta, function(s) {s[, 2]})
kurts <- c(mean(sapply(best2.semnig, kurtosis), na.rm=TRUE), 
           mean(sapply(best1.lasso, kurtosis), na.rm=TRUE),
           mean(sapply(best1.ridge, kurtosis), na.rm=TRUE), 
           mean(sapply(best1.bSEM, kurtosis), na.rm=TRUE))
ranges <- rbind(range(unlist(lapply(1:length(
  fit2.semnig$eb$mpriord), function(d) {
    fit2.semnig$eb$mpriord[d]*fit2.semnig$eb$mpriorf[[d]]}))),
  fit1.bSEM$eb$b[nrow(fit1.bSEM$eb$b), ]/
    (fit1.bSEM$eb$a[nrow(fit1.bSEM$eb$a), ] - 1))
@
  The non-zero $\bm{\alpha}$ estimates in Tables 
  \ref{tab:analysis_gdsc_est1_feat} and \ref{tab:analysis_gdsc_est1_drug} show 
  that there is an effect of the
  external covariates. Models NIG$_{\text{f}+\text{d}}$, NIG$_{\text{f}}$, and
  bSEM estimate a negative effect for
  the external pathway covariate. This translates to a negative additive effect 
  on the expected prior precision of the $\bm{\beta}_d$, i.e., genes that are 
  found
  in the target pathway of a drug are more predictive of IC50 than genes that
  are not in the target pathway, according to expectation. Furthermore the 
  NIG$_{\text{f}+\text{d}}$
  model estimates positive effects for the experimental and developmental drugs,
  compared to the approved drugs. This translates to smaller expected 
  prior precisions for the regression parameters of approved drugs, i.e.,
  approved drugs are easiest to predict, followed by developmental and 
  experimental drugs, as expected. On the other hand, the NIG$_{\text{d}}$ 
  model, that only includes external
  drug covariates, estimates a negative effect for the developmental drugs.
  The NIG$_{\text{f}+\text{d}}$ model estimates the smallest prior 
  $\bm{\beta}_d$ precision for the
  targeted drugs, followed by cytotoxic drugs, while drugs with an unkown target
  are assigned the largest prior $\bm{\beta}_d$ precision, according
  to expectation. Somewhat surprisingly, the NIG$_{\text{d}}$ model
  estimates the smallest
  prior $\bm{\beta}_d$ precision for the cytotoxic drugs, compared to the 
  targeted and unkown drugs. We conjecture that, due to absence of a local
  feature-specific variance component, this model is not flexible 
  enough to correctly capture the full covariance structure of the 
  $\bm{\beta}_d$ and therefore incorreclty estimates $\bm{\alpha}$.
  To see the total effect of the external covariates on the $\bm{\beta}_d$ 
  prior, we calculated the range of the expected prior variances 
  (up to a multiplicative error variance) as 
  $\hat{E}(\gamma_{jd}^2 \cdot \tau_d^2) \in \[ \Sexpr{ranges[1, ]}\]$ and
  $\hat{E}(\gamma_{jd}^2 \cdot \tau_d^2) \in \[ \Sexpr{ranges[2, ]}\]$ for
  NIG$_{\text{f}+\text{d}}$ and bSEM, respectively. This confirms that the
  estimated prior 
  $\bm{\beta}_d$ variance components are in the same order of magnitude and that
  the external covariates have a detectable effect on the prior.
<<analysis_gdsc_est1_feat>>=
options(knitr.kable.NA="", digits=3)
kable(est1,
      col.names=colnames(est1), align="r",
      caption="$\\bm{\\alpha}_{\\text{feat}}$ estimates from GDSC data.",
      format="latex", booktabs=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("HOLD_position"))
@

<<analysis_gdsc_est1_drug>>=
kable(est2,
      col.names=colnames(est2), align="r",
      caption="$\\bm{\\alpha}_{\\text{drug}}$ estimates from GDSC data.",
      format="latex", booktabs=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("HOLD_position"))
@
  The median PMSEs and ELBOs, calculated on the test data
  are displayed in Table \ref{tab:analysis_gdsc_tab1}. In terms of prediction 
  error, ridge outperforms the other models 
  (PMSE of \Sexpr{round(tab[6, 1], 3)}), while bSEM 
  (PMSE of \Sexpr{round(tab[7, 1], 3)}) performs worse than 
  the null model (PMSE of \Sexpr{round(tab[8, 1], 3)}). Second and third best
  performing models are NIG$_{\text{f}+\text{d}}^-$ and 
  NIG$_{\text{f}+\text{d}}$ (PMSEs of \Sexpr{round(tab[1, 1], 3)} and
  \Sexpr{round(tab[2, 1], 3)}, respectively), both outperforming the lasso
  with a PMSE of \Sexpr{round(tab[5, 1], 3)}. The underperformance of the 
  NIG$_{\text{f}}$ and NIG$_{\text{d}}$ models shows the usefulness of a
  global-local shrinkage prior as opposed to only global or local shrinkage as 
  as in models NIG$_{\text{f}}$ and NIG$_{\text{d}}$.
<<analysis_gdsc_tab1>>=
kable(tabm,
      col.names=colnames(tabm), align="r",
      caption="Performance measures calculated on GDSC data (best performing model in bold).",
      format="latex", booktabs=TRUE, escape=FALSE) %>%
  kable_styling(latex_options=c("HOLD_position"))
@

@@

	Simulations in Section \ref{sec:simulations} show that the method is 
	reasonably able to estimate the hyperparameters. Simulation Scenarios 1 and 2
	show
	that estimation of drug- and feature-specific hyperparameters is, in 
	principle, fairly accurate. However, when estimated jointly, biases may occur
	due to the interplay between the two sources of information. Nevertheless,
	the simulations show that inclusion of informative external covariates is 
	still beneficial to predictive performance.

	The model is put into practice on the GDSC data. 
	The NIG model is competitive with convential models like lasso in 
	terms of predictive performance, but is slightly outperformed by ridge.
	We note that estimation of the penalty parameters in the latter two models
	is geared towards predictive performance, because cross-validation is used
	to train these parameters, as opposed to empirical Bayes estimation of the
	hyperparameters in the NIG model, which aims to improve model fit. In that 
	sense, the comparison metric (PMSE) is not to the benefit of the NIG model.

@@

	A possible direction of future research to attempt
	to lower PMSE of the NIG model is the introduction of a cross-validated global
	variance component in the prior of the $\bm{\beta}_d$. If we add the 
	additional constraint that the drug and feature-specific variance components
	average to one, this cross-validated variance component may be interpreted as
	the overall level of shrinkage. This overall shrinkage level is then tailored 
	to lower PMSE, while the feature- and drug-specific variance component 
	include the external information. Other directions of future research are
	applications of the NIG
	model to different data types. One possibility is to apply the model in an 
	eQTL study, 
	where gene expressions are regressed on SNPs. Several interesting external
	covariates are available, both on the genes as well as the SNPs. An example of
	an external
	covariate for the genes is genelength, where we suspect that longer genes are
	harder to predict. The distance of the SNP to the gene is an example of an
	external SNP covariate, where the expectation is that SNPs further from the 
	gene are less predictive of that gene's expression.
